Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Remove duplicated proof
Improve text

\begin{corollary}
\label{C1}
Possibly, a God-like being exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} \hfill
\begin{prooftree}
\AXC{Axiom \ref{A3}} \dashedLine
\UIC{$P(G)$}
\AXC{Theorem \ref{T1} for G} \dashedLine
\UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
\BIC{$\pos \ex x. G(x)$}
\end{prooftree}
\end{proof}

@@

Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle relies only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound. There have been many discussions of what modal axioms are needed for the ontological argument. In their manuscripts, G\"odel and Scott do not write explicitly which logic they use. However, a proof step in their notes clearly relies on {\bf S5}'s \emph{iteration principle}.
Anderson \citep[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}. Sobel \citep{sobel2}[p. 152] acknowledges Anderson's conjecture, claims that it is correct and tries to informally explain how his formal proofs could be modified to rely on {\bf B} instead of {\bf S5}. Nevertheless, neither Anderson nor Sobel presented any formal proof relying only on {\bf B}. The new formal proof presented in Section \ref{sec:newproof} relies on {\bf KB}, which is weaker than {\bf B}.

Curiously, in the new proof of Theorem 3, the actual existence of a God-like being ($\ex x. G(x)$) is proven \emph{before} its necessity ($\nec \ex x. G(x)$). This contrasts with G\"odel's proof, where the actual existence is never actually proven, although it can be derived as a trivial corollary of its necessity by applying the modal axiom T.

>>>>

\begin{corollary}
\label{C1}
Possibly, a God-like being exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} The proof of this corollary is identical to the proof of the same corollary shown in Section \ref{sec:scott}.
% \begin{prooftree}
% \AXC{Axiom \ref{A3}} \dashedLine
% \UIC{$P(G)$}
% \AXC{Theorem \ref{T1} for G} \dashedLine
% \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
% \BIC{$\pos \ex x. G(x)$}
% \end{prooftree}
\end{proof}

@@

Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle requires only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound. There have been many discussions of what modal axioms are needed for the ontological argument. In their manuscripts, G\"odel and Scott do not write explicitly which logic they use. However, a proof step in their notes clearly relies on {\bf S5}'s \emph{iteration principle}.
Anderson \citep[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}. Sobel \citep{sobel2}[p. 152] acknowledges Anderson's conjecture, claims that it is correct and tries to informally explain how his formal proofs could be modified to rely on {\bf B} instead of {\bf S5}. Nevertheless, neither Anderson nor Sobel presented any formal proof relying only on {\bf B}. The new formal proof presented in this section relies on {\bf KB}, which is weaker than {\bf B}.

Curiously, in the new proof of Theorem 3, the actual existence of a God-like being ($\ex x. G(x)$) is proven \emph{before} its necessity ($\nec \ex x. G(x)$). This contrasts with G\"odel's proof, where the actual existence is never actually proven. It could, nevertheless, be derived as a trivial corollary by applying the modal axiom T.
